---
title: "2018_TOC"
author: "Melia Hodson"
date: "11/3/2020"
output: github_document
---

# Intro

This document shows how **individual bottle** TOC/DOC data from 2018 remineralization experiment was processed, quality controlled, and analyzed. 

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
```

# Import Data

```{r Retreiving Data, message=FALSE, warning=FALSE}
excel_sheets("~/Desktop/Github/144L_demo/stuff/Input_Data/week4/144L_2018_Exp_TOC.xlsx")

metadata <- read_excel("~/Desktop/Github/144L_demo/stuff/Input_Data/week4/144L_2018_Exp_TOC.xlsx", sheet = "Metadata")

data <- read_excel("~/Desktop/Github/144L_demo/stuff/Input_Data/week4/144L_2018_Exp_TOC.xlsx", sheet = "Data")
```

```{r metadata}
glimpse(metadata)
```

```{r data}
glimpse(data)
```

# Prepare Data

```{r}
joined <- left_join(metadata,data)

toc <- joined %>% 
  mutate(Datetime = ymd_hm(Datetime)) %>% 
  group_by(Bottle) %>% 
  mutate(interv = interval(first(Datetime), Datetime),
         hours = as.numeric(interv)/3600,
         days = hours/24) %>% 
  rename(sd_TOC = TOC_sd,
         sd_PTOC = PTOC_sd) %>% 
  ungroup() %>% 
  select(Experiment:Datetime, hours, days, everything(), -c(contains("Sample")), interv)

```

```{r}
glimpse(toc)
```

# Plot the curves

## Pivot Data

```{r}
pivot_toc <- toc %>% 
  select(Bottle, Treatment, days, TOC, PTOC) %>% 
  pivot_longer(TOC:PTOC, names_to = "sample", values_to = "values")

pivot_toc_sd <- toc %>% 
  select(Bottle, Treatment, days, sd_TOC, sd_PTOC) %>% 
  pivot_longer(sd_TOC:sd_PTOC, names_to = "sample", names_prefix = "sd_", values_to = "sd")

pivoted <- left_join(pivot_toc,pivot_toc_sd) %>% 
  mutate(sample = ifelse(sample == "TOC", "Bottle", "Vial"))

glimpse(pivoted)
```

## Plot

```{r fig.height=7, fig.width=8}
custom.colors <- c("Control" = "black", "Ash Leachate" = "royalblue3", "Mud Leachate" = "lawngreen", "Glucose_Nitrate_Phosphate" = "tan1", "Vial" = "hotpink2", "Bottle" = "forestgreen")

levels <- c("Control", "Ash Leachate", "Mud Leachate", "Glucose_Nitrate_Phosphate", "Bottle", "Vial")

pivoted %>% 
  drop_na(values) %>% 
  mutate(Treatment = factor(Treatment, levels = levels),
         sample = factor(sample, levels)) %>% 
  ggplot(aes(x = days, y = values, group = interaction(Treatment, Bottle))) +
  geom_errorbar(aes(ymin = values - sd, ymax = values + sd, color = sample), width = 0.4) +
  geom_point(aes(fill = sample), size = 3, shape = 21, alpha = 0.7) +
  scale_color_manual(values = custom.colors) +
  scale_fill_manual(values = custom.colors) +
  labs(x = "Days", y = expression("Total Organic Carbon, umol C L"^-1), color = "", fill = "") +
  theme_bw() +
  facet_grid(rows = "Treatment", scales = "free") +
  guides(color = F) +
  ggtitle("TOC and PTOC Treatment Values and \nStandard Deviations")
```

The analytical detection limit for our TOC analysis is ~1.5 umol C L^-1^.

# TOC v PTOC (bottle v vial samples)

```{r message=FALSE, warning=FALSE}
# install.packages("lmodel2")
library(lmodel2)

reg.data <- toc %>% 
  drop_na(TOC) %>% 
  filter(Timepoint > 0) 

reg <- lmodel2(PTOC ~ TOC, data = reg.data, nperm = 99) 
```

```{r}
reg
```
```{r}
intercept <- reg$regression.results[3,2]
slope <- reg$regression.results[3,3]

two_int <- reg$confidence.intervals[3,2]
two_slope <- reg$confidence.intervals[3,4]
nine_int <- reg$confidence.intervals[3,3]
nine_slope <- reg$confidence.intervals[3,5]
```

```{r}
reg.data %>% 
  ggplot(aes(x = TOC, y = PTOC)) +
  geom_errorbar(aes(ymin = PTOC - sd_PTOC, ymax = PTOC + sd_PTOC), width = 0.05) +
  geom_point(shape = 21, size = 4, alpha = 0.7) +
  geom_abline(intercept = intercept, slope = slope, linetype = 2, size = 1) +
  geom_abline(intercept = two_int, slope = two_slope, color = "black", linetype = 3, size = 1) +
  geom_abline(intercept = nine_int, slope = nine_slope, color = "black", linetype = 3, size = 1) +
  labs(x = expression("Bottle TOC, umol C L"^-1), y = expression("Vial TOC umol C L"^-1)) +
  theme_bw() +
  annotate(geom = "text", label = expression(atop("y = 1.03x - 2.12", paste("r"^2, "= 0.922", italic(", p"), "= 0.01"))), x = 81, y = 88, size = 4)
```
This graph looks at the relative concentrations of TOC in the bottle samples versus the vial samples. There appears to be a linear relationship between the two.

```{r}
bc <- read_rds("~/Desktop/Github/144L_demo/stuff/Output_Data/Week 3/2018_Processed_Experiment_Data.rds")

merge <- left_join(toc, bc) %>% 
  select(Experiment:days, TOC:sd_PTOC, cells:diff_ln_cells, bc, ave_bc, sd_bc, everything(), - c(contains("Sample"))) %>% 
  group_by(Treatment, Bottle) %>% 
  fill(exp_start:sd_lag) %>% 
  ungroup()

glimpse(merge)

subset <- merge %>% 
  select(Experiment, Treatment, Bottle, days, PTOC, bc)
```

# Interpolate

```{r message=FALSE, warning=FALSE}
# install.packages("zoo")
library(zoo)
```

We only want to interpolate within an experiment and not across experiments, so we'll perform the interpolation by using the split - apply - combine strategy
- Split the dataframe into a list of its elements(in this case treatment and bottles)
- Apply the dataframe into a list of its elements
- Combine the results into a new dataframe

## Split

```{r}
to_interpolate <- merge %>% 
  select(Experiment, Bottle, Timepoint, days, PTOC, bc) %>% 
  group_by(Bottle)

list <- to_interpolate %>% 
  group_split()

keys <- to_interpolate %>% 
  group_keys() %>% 
  mutate(key = paste(Bottle))

names(list) <- keys$key
```

## Write the Function

```{r}
interp.fxn <- function(x) {
  y <- zoo(x, order.by = z$days)
  interp_toc <- round(as.numeric(na.approx(y$PTOC, na.rm = F)), 1)
  interp_bc <- round(as.numeric(na.approx(y$bc, na.rm = F)), 1)
  z <- cbind(y, interp_toc, interp_bc)
  as_tibble(z)}

glimpse(interp.fxn)
```

## Apply and Combine

```{r}
interpolated <- lapply(list, interp.fxn) %>% 
  plyr::ldyply(., as.data.frame) %>% 
  select(-c(.id, PTOC, bc, days)) %>% 
  mutate_at(vars(Timepoint:interp_bc), as.numeric) %>% 
  left_join(merge, .)

glimpse(interpolated)
```

# Estimate DOC, Bioavailablity, BGE's

```{r}
doc <- interpolated %>% 
  mutate(doc = interp_toc - interp_bc) %>% 
  group_by(Bottle) %>% 
  mutate(bioav_doc = first(doc) - last(na.omit(doc)) / first(doc),
         doc_exp_end = ifelse(Timepoint == exp_end, doc, NA),
         delta_doc = first(doc) - doc_exp_end,
         tdelta_doc = first(doc) - last(na.omit(doc)),
         bge = ifelse(delta_doc > 1.5, dellta_bc / delta_doc, NA)) %>% 
  fill(doc_exp_end:bge, .direction = "downup") %>% 
  ungroup()

glimpse(doc)
```

# Treatment Averages

```{r}
averages <- doc %>% 
  group_by(Treatment, Timepoint) %>% 
  mutate(ave_toc = mean(PTOC),
         sd_toc = sd(PTOC)) %>% 
  ungroup() %>% 
  group_by(Treatment) %>% 
  mutate(ave_bioav_doc = mean(bioav_doc),
         sd_bioav_doc = sd(bioav_doc),
         ave_delta_doc = mean(delta_doc),
         sd_delta_doc = sd(delta_doc),
         ave_tdelta_doc = mean(tdelta_doc),
         sd_tdelta_doc = sd(tdelta_doc),
         ave_bge = mean(bge),
         sd_bge = sd(bge)) %>% 
  ungroup()
```

# Plot Treatment Curves

```{r}
averages %>% 
  mutate(Treatment = factor(Treatment, levels = levels)) %>% 
  drop_na(ave_toc) %>% 
  ggplot(aes(x = days, y = ave_toc, group = interaction(Treatment))) +
  geom_errorbar(aes(ymin = ave_toc - sd_toc, ymax = ave_toc + sd_toc, color = Treatment), width = 0.4) +
  geom_point(aes(fill = Treatment), size = 3, shape = 21, alpha = 0.7) +
  geom_line(aes(color = Treatment), size = 1, alpha = 0.7) +
  scale_color_manual(values = custom.colors) +
  scale_fill_manual(values = custom.colors) +
  labs(x = "Days", y = expression("Total Organic Carbon, umol C L"^-1), color = "", fill = "") +
  theme_bw() +
  facet_grid(rows = "Treatment", scales = "free") +
  guides(color = F) +
  ggtitle("Treatment Averages Dependent and Independent /nof Timepoint")
```

# Plot Treatment Averages (Bars)

```{r}
bar.data <- averages %>% 
  select(Treatment, ave_bioav_doc:sd_bge) %>% 
  distinct()
```

```{r}
bioav <- bar.data %>% 
  ggpolt(aes(x = factor(Treatment, levels = levels), y = ave_bioav_doc, group = interaction(Treatment))) +
  geom_col(color = "black", fill = "white", alpha = 0.7) +
  geom_errorbar(aes(ymin = ave_bioav_doc - sd_bioav_doc, ymax = ave_bioav_doc + sd_bioav_doc), width = 0.1, size = 0.5) +
  labs(x = "", y = expression(paste("Bioavailible DOC Fraction")), color = "") +
  theme_bw()
```

```{r}
library(patchwork)
```

```{r}
del_doc / total_doc / bioav / bge + plot_annotation(title = "2018 144L Experiments", tag_levels = "a")
```

# Save Data

```{r}
saveRDS(averages, "~/Desktop/Github/144L_demo/stuff/Output_Data/Week 4/2018_Experiment_Processed_DOC_BGE.Rmd")
```

