---
title: "EEMB 144L 2018 Experiment Bacterial Abundance"
author: "Melia Hodson"
date: "10/30/2020"
output: github_document
---
# Intro
This document shows how *Individual Bottle* bacterial abundance data from the 2018 EEMB 144L experiments was processed, quality controlled, and analyzed. 

```{r Loading Packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
library(RColorBrewer)
```

# Import Data

```{r Data for 2018 Experiments}
excel_sheets("~/Desktop/Github/144L_demo/stuff/Input_Data/week3/144L_2018_BactAbund.xlsx")

metadata <- read_excel("~/Desktop/Github/144L_demo/stuff/Input_Data/week3/144L_2018_BactAbund.xlsx", sheet = "Metadata")
unique(metadata$Experiment)
unique(metadata$Location)
unique(metadata$Bottle)
unique(metadata$Treatment)

data <- read_excel("~/Desktop/Github/144L_demo/stuff/Input_Data/week3/144L_2018_BactAbund.xlsx", sheet = "Data")
glimpse(data)

joined <- left_join(metadata,data)
names(joined)
summary(joined)
glimpse(joined)
```

# Prepare Data

Convert date and time column values from characters to date, add columns with time elapsed during the experiment, convert cells/mL to cells/L, and subset data to select only for variables of interest and drop NA values.

```{r Data Preparation, message=FALSE, warning=FALSE}
cells <- joined %>% 
  mutate(Datetime = ymd_hm(Datetime), 
         cells = Cells_ml * 1000) %>% 
  group_by(Treatment, Bottle) %>% 
  mutate(interv = interval(first(Datetime), Datetime),
         hours = as.numeric(interv)/3600,
         days = hours/24) %>% 
  ungroup() %>% 
  select(Experiment:DNA_Sample, hours, days, cells) %>% 
  drop_na()
  
glimpse(cells)
```

# Plot Growth Curves

```{r Growth Curves for the Data Subsets}
custom.colors <- c("Control" = "black", "Ash Leachate" = "seagreen3", "Mud Leachate" = "dodgerblue", "Glucose_Nitrate_Phosphate" = "darkblue")

levels <- c("Control", "Ash Leachate", "Mud Leachate", "Glucose_Nitrate_Phosphate")

cells %>% 
  mutate(dna = ifelse(DNA_Sample == T, "*",NA)) %>% 
  ggplot(aes(x = days, y = cells, group = interaction(Treatment, Bottle))) +
  geom_line(aes(color = factor(Treatment, levels =levels)),size = 1) +
geom_point(aes(fill = Treatment), size = 3, color = "black", shape = 21) +
  geom_text(aes(label = dna), size = 12, color = "red") +
  labs(x = "Days", y = expression(paste("Cells, L"^-1)), fill = "") +
  guides(color = F) +
  scale_color_manual(values = custom.colors) +
  scale_fill_manual(values = custom.colors) +
  theme_bw()


subset <- cells %>% select(Treatment, Bottle, days, cells)
subset <- cells
glimpse(subset)
```
We can calculate:
- The total change in cells from the initial condition to the end of the experiment
- Specific growth rates as the slope of ln(abundance) v time during the exponential growth phase
- Doubling time as ln(2) divided by the specific growth rate
- The mean of each of these parameters for each treatment

First we'll need to determine where exponential growth occurs in the experiment, if it does. So next we will plot ln(abundance) v time


# Identify Exponential Phase of Growth

```{r Exponential Possibility}
ln_cells <- cells %>% 
  group_by(Treatment, Bottle) %>% 
  mutate(ln_cells = log(cells),
         diff_ln_cells = ln_cells - lag(ln_cells, default = first(ln_cells))) %>% 
  ungroup()
```


```{r Graphing, fig.height=4, fig.width=10, warning=FALSE}
ln_cells %>% 
  mutate(dna = ifelse(DNA_Sample == T, "*",NA)) %>% 
  ggplot(aes(x = days, y = diff_ln_cells, group = interaction(Treatment, Bottle))) +
  geom_line(aes(color = factor(Treatment, levels =levels)),size = 1) +
geom_point(aes(fill = Treatment), size = 3, color = "black", shape = 21) +
  geom_text(aes(label = dna), size = 12, color = "red") +
  labs(x = "Days", y = expression(paste("Change in ln of Cells, L"^-1)), fill = "") +
  guides(color = F) +
  scale_color_manual(values = custom.colors) +
  scale_fill_manual(values = custom.colors) +
  theme_bw() + 
  facet_grid(Location~Bottle, scales = "free")
```
This graph is showing separation of the different treatments among bottles A- H, and you can see exponential growth within the first two timepoints.

# Calculating growth rates, delta cell abundances, and doubling Time

- Bottle A ~1-2d (T2-T4)
- Bottle B ~0-2d (T0-T3)
- Bottle C ~0-1d (T0-T1)
- Bottle D ~0-1d (T0-T1)
- Bottle E ~0-1d (T0-T1)
- Bottle F ~0-1d (T0-T1)
- Bottle G ~1-2d (T1-T3)
- Bottle H ~1-2d (T1-T3)

```{r}
compare <- ln_cells %>% 
  select(Treatment, Bottle, Timepoint, days, diff_ln_cells)

growth <- ln_cells %>% 
  mutate(exp_start = 0,
         exp_end = ifelse(!Bottle %in% c("G", "H"), 1, 2)) %>% 
  group_by(Treatment, Bottle) %>% 
  mutate(ln_cells_exp_start = ifelse(Timepoint == exp_start, ln_cells, NA),
         ln_cells_exp_end = ifelse(Timepoint == exp_end, ln_cells, NA),
         cells_exp_start = ifelse(Timepoint == exp_start, cells, NA),
         cells_exp_end = ifelse(Timepoint == exp_end, cells, NA),
         days_exp_start = ifelse(Timepoint == exp_start, days, NA),
         days_exp_end = ifelse(Timepoint == exp_end, days, NA),) %>% 
  fill(ln_cells_exp_start:days_exp_end, .direction = "updown") %>% 
  mutate(mu = (ln_cells_exp_end - ln_cells_exp_start)/(days_exp_end - days_exp_start),
         doubling = log(2)/mu,
         delta_cells = cells_exp_end - first(cells)) %>% 
  ungroup()


glimpse(growth)
``` 

# Convert bacterial abundance and change in bacterial abundance to carbon units 

Apply a carbon conversion factor (CCF) to bacterial abundance (cells L^-1^) to generate bacterial carbon (umol C L^-1^)

We'll apply the average carbon content of bacterioplankton cells from Coastal Japan (~30 fg C cell ^-1^), reported by [Fukuda et al., 1998]

This CCF was used in previous experiments conducted in the SBC: [James et al., 2017]

```{r}
bactcarbon <- growth %>% 
  mutate(bc = cells * (2.5 * 10^-9),
         delta_bc = delta_cells * (2.5 * 10^-9))
glimpse(bactcarbon)
```

# Calculating Treatment Averages

```{r}
averages <- bactcarbon %>% 
  group_by(Treatment, Timepoint) %>% 
  mutate(ave_bc = mean(bc),
         sd_bc = sd(bc)) %>% 
  ungroup() %>% 
  group_by(Treatment) %>% 
  mutate(ave_mu = mean(mu),
         sd_mu = sd(mu),
        ave_doubling = mean(doubling),
        sd_doubling = sd(doubling),
        ave_delta_cells = mean(delta_cells),
        sd_delta_cells = sd(delta_cells),
        ave_delta_bc = mean(delta_bc),
        sd_delta_bc = sd(delta_bc),
        ave_lag = mean(days_exp_start),
        sd_lag = sd(days_exp_start)) %>% 
  ungroup()

glimpse(averages)

subset <- averages %>% 
  select(Treatment, Bottle, Timepoint, bc, ave_bc, sd_bc)
```

# Plot Treatmnt Averages

```{r}
averages %>% 
  ggplot(aes(x = days, y = ave_bc), group = interaction(Treatment))  + 
  geom_errorbar(aes(ymin = ave_bc - sd_bc, ymax= ave_bc + sd_bc, color = factor(Treatment, levels = levels)), width = 0.1) +
  geom_line(aes(color = factor(Treatment, levels = levels)), size = 1) +
  geom_point(aes(fill = factor(Treatment,levels =levels)), color = "black", shape = 21, size = 3) +
   scale_color_manual(values = custom.colors) +
  scale_fill_manual(values = custom.colors, name = "Treatments") +
  theme_bw() +
  labs(x = "Days", y = expression(paste("Bacterial Carbon, umol C L"^-1), fill = "", color = "")) +
    guides(color = F) +
  ggtitle("Bacterial Abundance Treatment Averages")
```

## Bar Plots

```{r}
bar.data <- averages %>% 
  select(Location, Treatment, ave_mu:sd_lag) %>% 
  distinct()

glimpse(bar.data)
```

```{r}
mu <- bar.data %>% 
  ggplot(aes(x = factor(Treatment, levels = levels), y = ave_mu), group = interaction(Treatment)) +
  geom_col(color = "black", fill = "blue") +
  geom_errorbar(aes(ymin = ave_mu - sd_mu, ymax = ave_mu + sd_mu), width = 0.1) +
  labs(x = "", y = expression("u, d"^-1)) +
  theme_bw()
```

```{r}
doubling <- bar.data %>% 
  ggplot(aes(x = factor(Treatment, levels = levels), y = ave_doubling), group = interaction(Treatment)) +
  geom_col(color = "black", fill = "blue") +
  geom_errorbar(aes(ymin = ave_doubling - sd_doubling, ymax = ave_doubling + sd_doubling), width = 0.1) +
  labs(x = "", y = expression("Doubling Time, d"^-1)) +
  theme_bw()
```

```{r}
delta_bc <- bar.data %>% 
  ggplot(aes(x = factor(Treatment, levels = levels), y = ave_delta_bc), group = interaction(Treatment)) +
  geom_col(color = "black", fill = "blue") +
  geom_errorbar(aes(ymin = ave_delta_bc - sd_delta_bc, ymax = ave_delta_bc + sd_delta_bc), width = 0.1) +
  labs(x = "", y = expression("Delta Bacterial Carbon, umol C L"^-1)) +
  theme_bw()
```

```{r}
lag <- bar.data %>% 
  ggplot(aes(x = factor(Treatment, levels = levels), y = ave_lag), group = interaction(Treatment)) +
  geom_col(color = "black", fill = "blue") +
  geom_errorbar(aes(ymin = ave_lag - sd_lag, ymax = ave_lag + sd_lag), width = 0.1) +
  labs(x = "", y = "Lag Phase, days") +
  theme_bw()
```

```{r}
# install.packages("patchwork")
library(patchwork)
```

```{r}
delta_bc + mu + doubling + plot_annotation(tag_levels = "a")
```

# Saving our Data

```{r}
saveRDS(averages, "~/Desktop/Github/144L_demo/stuff/Output_Data/Week 3/2018_Processed_Experiment_Data.rds")
```

Test

